---
title: "AIPW: Augmented Inverse Probability Weighting"
output:
  github_document
---
<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/yqzhong7/AIPW/branch/master/graph/badge.svg)](https://codecov.io/gh/yqzhong7/AIPW?branch=master)
[![Travis build status](https://travis-ci.com/yqzhong7/AIPW.svg?branch=master)](https://travis-ci.com/yqzhong7/AIPW)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#WIP)
[![R build status](https://github.com/yqzhong7/AIPW/workflows/R-CMD-check/badge.svg)](https://github.com/yqzhong7/AIPW/actions)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

__Authors:__ [Yongqi Zhong](https://github.com/yqzhong7), [Ashley Naimi](https://github.com/ainaimi), [Gabriel Conzuelo](https://github.com/gconzuelo)


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4,
  fig.path = "man/figures/",
  comment = "#>"
)
```

## Contents:

  * ##### [Installation](#Installation)
  * ##### [Example](#Example)
    + ###### [Setup example data](#data)
    + ###### [One line version](#one_line)
    + ###### [Longer version](#details)
      1. [Create an AIPW object](#constructor)
      2. [Fit the object](#fit)
      3. [Calculate average treatment effects](#ate)
      4. [Parallelization](#par)


---------

## <a id="Installation"></a>Installation

```{r installation, eval=FALSE}
install.packages("remotes")
remotes::install_github("yqzhong7/AIPW")
```


## <a id="Example"></a>Example

### <a id="data"></a>Setup example data
```{r example data}
set.seed(888)
N <- 200
outcome <- rbinom(N,1,0.3)
exposure <- rbinom(N,1,0.2)
#covaraites for outcome model (Q)
covariates.Q <- matrix(c(rbinom(N,1,0.4),
                                rnorm(N,mean = 0,sd=1),
                                rpois(N,lambda = 2)),
                              ncol=3)
#covariates for exposure model (g)
covariates.g <- matrix(c(rbinom(N,1,0.4),
                                rnorm(N,mean = 0,sd=1),
                                rpois(N,lambda = 2)),
                              ncol=3)

# covariates.g <- c(rbinom(N,1,0.4)) #a vector of a single covariate is also supported
```

### <a id="one_line"></a>One line version (Method chaining from R6class)

```{r one_line, fig.width=8}
library(AIPW)
library(SuperLearner)
library(ggplot2)
library(progressr) #for the progress bar
AIPW_SL <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=FALSE)$
  fit()$
  calculate_result(g.bound = 0.25)$
  plot.p_score()
```


### <a id="details"></a>A slightly longer version

#### 1. <a id="constructor"></a>Create an AIPW object 

  * ##### Use [SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/index.html) libraries (reference: [Guide to SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/vignettes/Guide-to-SuperLearner.html))
```{r SuperLearner, message=FALSE}
library(AIPW)
library(SuperLearner)

#SuperLearner libraries for outcome (Q) and exposure models (g)
sl.lib <- c("SL.mean","SL.glm")

#construct an aipw object for later estimations 
AIPW_SL <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = sl.lib,
                    g.SL.library = sl.lib,
                    k_split = 3,
                    verbose=FALSE)
```

  * ##### Use [sl3](https://tlverse.org/sl3/index.html) libraries (reference: [Intro to sl3](https://tlverse.org/sl3/articles/intro_sl3.html))
```{r sl3}
library(AIPW)
library(sl3)

##construct sl3 learners for outcome (Q) and exposure models (g)
lrnr_glm <- Lrnr_glm$new()
lrnr_mean <- Lrnr_mean$new()
#stacking two learner (this will yield estimates for each learner)
stacklearner <- Stack$new(lrnr_glm, lrnr_mean) 
#metalearner is required to combine the estimates from stacklearner
metalearner <- Lrnr_nnls$new()
sl3.lib <- Lrnr_sl$new(learners = stacklearner,
                      metalearner = metalearner)

#construct an aipw object for later estimations 
AIPW_sl3 <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = sl3.lib,
                    g.SL.library = sl3.lib,
                    k_split = 3,
                    verbose=FALSE)
```

#### 2. <a id="fit"></a>Fit the AIPW object

This step will fit the data stored in the AIPW object to obtain estimates for later average treatment effect calculations.

```{r}
#fit the AIPW_SL object
AIPW_SL$fit()
# #fit the AIPW_sl3 object
# AIPW_sl3$fit()
```

#### 3. <a id="ate"></a>Calculate average treatment effects 

  * ##### Check the balance of propensity scores by exposure status
```{r ps_raw}
library(ggplot2)
AIPW_SL$plot.p_score()
```

  * ##### Estimate the ATE with propensity scores truncation
```{r}
#estimate the average causal effects from the fitted AIPW_SL object 
AIPW_SL$calculate_result(g.bound = 0.25) #propensity score truncation 

#estimate the average causal effects from the fitted AIPW_sl3 object 
# AIPW_sl3$calculate_result(g.bound = 0.25) #propensity score truncation 
```

  * ##### Check the balance of propensity scores by exposure status after truncation
```{r ps_trunc, fig.width=8}
AIPW_SL$plot.p_score()
```

#### 4. <a id="par"></a>Parallelization with future.apply

The current version of AIPW package supports parallel processing using [future.apply](https://github.com/HenrikBengtsson/future.apply) package under the [future](https://github.com/HenrikBengtsson/future) framework. Simply use `plan()` to enable this feature:

```{r parallel, eval=F}
# install.packages("future.apply")
library(future.apply)
plan(multiprocess, workers=5, gc=T)
AIPW_SL <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = sl.lib,
                    g.SL.library = sl.lib,
                    k_split = 3,
                    verbose=F)$fit()$calculate_result()
```
