---
title: "AIPW: Augmented Inverse Probability Weighting"
output:
  github_document
---
<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Codecov test coverage](https://codecov.io/gh/yqzhong7/AIPW/branch/master/graph/badge.svg)](https://codecov.io/gh/yqzhong7/AIPW?branch=master)
[![Travis build status](https://travis-ci.com/yqzhong7/AIPW.svg?branch=master)](https://travis-ci.com/yqzhong7/AIPW)
[![R build status](https://github.com/yqzhong7/AIPW/workflows/R-CMD-check/badge.svg)](https://github.com/yqzhong7/AIPW/actions)

<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

__Authors:__ [Yongqi Zhong](https://github.com/yqzhong7), [Ashley Naimi](https://github.com/ainaimi), [Gabriel Conzuelo](https://github.com/gconzuelo), [Edward Kennedy](https://github.com/ehkennedy)


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4,
  fig.path = "man/figures/",
  comment = "#>"
)
```

----------------

Augmented inverse probability weighting (AIPW) is a doubly robust estimator for causal inference. The `AIPW` package is designed for estimating the average treatment effect of a binary exposure and outcome on risk difference (RD), risk ratio (RR) and odds ratio (OR). Users need to examine causal assumptions (e.g., consistency) before using this package. 

---------------

## Contents:

  * ##### [Installation](#Installation)
  * ##### [Example](#Example)
    + ###### [Setup example data](#data)
    + ###### [One line version](#one_line)
  * ##### [Parallelization and progress bar](#par) 
  * ##### [Use tmle/tmle3 as input](#tmle)
  * ##### [References](#ref)


---------

## <a id="Installation"></a>Installation

```{r installation, eval=FALSE}
install.packages("remotes")
remotes::install_github("yqzhong7/AIPW")
```


## <a id="Example"></a>Example

### <a id="data"></a>Setup example data
```{r example data}
set.seed(123)
data("eager_sim_obs")
outcome <- eager_sim_obs$sim_A
exposure <- eager_sim_obs$sim_Y
#covariates for both outcome model (Q) and exposure model (g)
covariates <- eager_sim_obs[-1:-2]

# covariates <- c(rbinom(N,1,0.4)) #a vector of a single covariate is also supported
```

### <a id="one_line"></a>One line version (`AIPW` class: method chaining from R6class)

```{r one_line, fig.width=8}
library(AIPW)
library(SuperLearner)
library(ggplot2)
AIPW_SL <- AIPW$new(Y = outcome,
                    A = exposure,
                    W = covariates, 
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=FALSE)$
  fit()$
  summary(g.bound = 0.25)$
  plot.p_score()
```

To see the results, set `verbose = TRUE`(default) or:
```{r res}
AIPW_SL$result
```

You can also use the `aipw_wrapper()` to wrap `new()`, `fit()` and `summary()` together (also support method chaining):
```{r wrapper, eval=F}
AIPW_SL <- aipw_wrapper(Y = outcome,
                        A = exposure,
                        W = covariates, 
                        Q.SL.library = c("SL.mean","SL.glm"),
                        g.SL.library = c("SL.mean","SL.glm"),
                        k_split = 3,
                        verbose=TRUE)$plot.p_score()
```

## <a id="par"></a>Parallelization with `future.apply` and progress bar with `progressr`

In default setting, the `AIPW$fit()` method will be run sequentially. The current version of AIPW package supports parallel processing implemented by [future.apply](https://github.com/HenrikBengtsson/future.apply) package under the [future](https://github.com/HenrikBengtsson/future) framework. Simply use `future::plan()` to enable parallelization and `set.seed()` to take care of the random number generation (RNG) problem:

```{r parallel, eval=T}
###Additional steps for parallel processing###
# install.packages("future.apply")
library(future.apply)
future::plan(multiprocess, workers=2, gc=T)
set.seed(888)

###Same procedure for AIPW as described above###
AIPW_SL <- AIPW$new(Y = outcome,
                    A = exposure,
                    W = covariates, 
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=TRUE)$fit()$summary()
```

Progress bar that supports parallel processing is available in the `AIPW$fit()` method through the API from [progressr](https://github.com/HenrikBengtsson/progressr) package:
```{r progressr, eval=F}
library(progressr)
#define the type of progress bar
handlers("progress")
#reporting through progressr::with_progress() which is embedded in the AIPW$fit() method
with_progress(
  AIPW_SL <- AIPW$new(Y = outcome,
                    A = exposure,
                    W = covariates, 
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=FALSE)$fit()$summary()
)
#also available for the wrapper
with_progress(
  AIPW_SL <- aipw_wrapper(Y = outcome,
                        A = exposure,
                        W = covariates, 
                        Q.SL.library = c("SL.mean","SL.glm"),
                        g.SL.library = c("SL.mean","SL.glm"),
                        k_split = 3,
                        verbose=FALSE)
)
```

## <a id="tmle"></a>Use `tmle`/`tmle3` fitted object as input (`AIPW_tmle` class)

`AIPW_tmle` class is designed for using `tmle`/`tmle3` fitted object as input

#### 1. `tmle`

```{r tmle}
require(tmle)
require(SuperLearner)
tmle_fit <- tmle(Y = outcome, A = exposure,W = covariates,
                 Q.SL.library=c("SL.mean","SL.glm"),
                 g.SL.library=c("SL.mean","SL.glm"),
                 family="binomial")
AIPW_tmle$
  new(A=exposure,Y=outcome,tmle_fit = tmle_fit,verbose = TRUE)$
  summary(g.bound=0.025)
```

#### 2. `tmle3`

```{r tmle3, eval =FALSE}
library(sl3)
library(tmle3)
node_list <- list(A = "sim_A",Y = "sim_Y",W = colnames(eager_sim_obs)[-1:-2])
or_spec <- tmle_OR(baseline_level = "0",contrast_level = "1")
tmle_task <- or_spec$make_tmle_task(eager_sim_obs,node_list)
lrnr_glm <- make_learner(Lrnr_glm)
lrnr_mean <- make_learner(Lrnr_mean)
sl <- Lrnr_sl$new(learners = list(lrnr_glm,lrnr_mean))
learner_list <- list(A = sl, Y = sl)
tmle3_fit <- tmle3(or_spec, data=eager_sim_obs, node_list, learner_list)

# parse tmle3_fit into AIPW_tmle class
AIPW_tmle$
  new(A=eager_sim_obs$sim_A,Y=eager_sim_obs$sim_Y,tmle_fit = tmle3_fit,verbose = TRUE)$
  summary()
```


--------------------
### <a id="ref"></a>References:

Robins, J.M., Rotnitzky, A. and Zhao, L.P., 1994. Estimation of regression coefficients when some regressors are not always observed. Journal of the American statistical Association, 89(427), pp.846-866.

Glynn, A.N. and Quinn, K.M., 2010. An introduction to the augmented inverse propensity weighted estimator. Political analysis, 18(1), pp.36-56.

Pearl, J., 2009. Causality. Cambridge university press.


