---
title: "AIPW: Augmented Inverse Probability Weighting"
output:
  github_document
---
<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#WIP)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/yqzhong7/AIPW/branch/master/graph/badge.svg)](https://codecov.io/gh/yqzhong7/AIPW?branch=master)
[![Travis build status](https://travis-ci.com/yqzhong7/AIPW.svg?branch=master)](https://travis-ci.com/yqzhong7/AIPW)
[![R build status](https://github.com/yqzhong7/AIPW/workflows/R-CMD-check/badge.svg)](https://github.com/yqzhong7/AIPW/actions)

<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

__Authors:__ [Yongqi Zhong](https://github.com/yqzhong7), [Ashley Naimi](https://github.com/ainaimi), [Gabriel Conzuelo](https://github.com/gconzuelo)


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4,
  fig.path = "man/figures/",
  comment = "#>"
)
```

## Contents:

  * ##### [Installation](#Installation)
  * ##### [Example](#Example)
    + ###### [Setup example data](#data)
    + ###### [One line version](#one_line)
  * ##### [Parallelization](#par)
  * ##### [Use tmle/tmle3 as input](#tmle)


---------

## <a id="Installation"></a>Installation

```{r installation, eval=FALSE}
install.packages("remotes")
remotes::install_github("yqzhong7/AIPW")
```


## <a id="Example"></a>Example

### <a id="data"></a>Setup example data
```{r example data}
set.seed(888)
N <- 200
outcome <- rbinom(N,1,0.4)
exposure <- rbinom(N,1,0.3)
#covariates for both outcome model (Q) and exposure model (g)
covariates <- matrix(c(rbinom(N,1,0.4),
                       rnorm(N,mean = 0,sd=1),
                       rpois(N,lambda = 2)),
                     ncol=3)

# covariates <- c(rbinom(N,1,0.4)) #a vector of a single covariate is also supported
```

### <a id="one_line"></a>One line version (`AIPW` class: method chaining from R6class)

```{r one_line, fig.width=8}
library(AIPW)
library(SuperLearner)
library(ggplot2)
library(progressr) #for the progress bar
AIPW_SL <- AIPW$new(Y = outcome,
                    A = exposure,
                    W = covariates, 
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=FALSE)$
  fit()$
  summary(g.bound = 0.25)$
  plot.p_score()
```

To see the results, use `verbose = TRUE` option or:
```{r res}
AIPW_SL$result
```

You can also use the `aipw_wrapper()` to wrap `new()`, `fit()` and `summary()` together (also support method chaining):
```{r wrapper, eval=F}
AIPW_SL <- aipw_wrapper(Y = outcome,
                        A = exposure,
                        W = covariates, 
                        Q.SL.library = c("SL.mean","SL.glm"),
                        g.SL.library = c("SL.mean","SL.glm"),
                        k_split = 3,
                        verbose=TRUE)$plot.p_score()
```

## <a id="par"></a>Parallelization with `future.apply`

In default setting, the `AIPW$fit()` method will be run sequentially. The current version of AIPW package supports parallel processing implemented by [future.apply](https://github.com/HenrikBengtsson/future.apply) package under the [future](https://github.com/HenrikBengtsson/future) framework. Simply use `future::plan()` to enable parallelization and `set.seed()` to take care of the random number generation (RNG) problem:

```{r parallel, eval=T}
# install.packages("future.apply")
library(future.apply)
future::plan(multiprocess, workers=2, gc=T)
set.seed(888)
AIPW_SL <- AIPW$new(Y = outcome,
                    A = exposure,
                    W = covariates, 
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=TRUE)$fit()$summary()
```

## <a id="tmle"></a>Use `tmle`/`tmle3` fitted object as input (`AIPW_tmle` class)

`AIPW_tmle` class is designed for using `tmle`/`tmle3` fitted object as input

#### 1. `tmle`

As shown in the message, [tmle](https://cran.r-project.org/web/packages/tmle/index.html) alone does not support sample splitting. In addition, `tmle` does not support different covariate sets for the exposure and the outcome models, respectively.

```{r tmle}
require(tmle)
require(SuperLearner)
tmle_fit <- tmle(Y = outcome, A = exposure,W = covariates,
                 Q.SL.library=c("SL.mean","SL.glm"),
                 g.SL.library=c("SL.mean","SL.glm"),
                 family="binomial")
AIPW_tmle$
  new(A=exposure,Y=outcome,tmle_fit = tmle_fit,verbose = TRUE)$
  summary(g.bound=0.025)
```

#### 2. `tmle3`

Similarly, [tmle3](https://github.com/tlverse/tmle3) may not support different covariate sets for the exposure and the outcome models, respectively. However, `tmle3` conducts sample splitting and propensity truncation (0.025) by default.

```{r tmle3}
require(tmle3)
require(sl3)
df <- data.frame(exposure,outcome,covariates)
node_list <- list(A = "exposure",Y = "outcome",W = colnames(df)[-1:-2])
or_spec <- tmle_OR(baseline_level = "0",contrast_level = "1")
tmle_task <- or_spec$make_tmle_task(df,node_list)
lrnr_glm <- make_learner(Lrnr_glm)
lrnr_mean <- make_learner(Lrnr_mean)
sl <- Lrnr_sl$new(learners = list(lrnr_glm,lrnr_mean))
learner_list <- list(A = sl, Y = sl)
tmle3_fit <- tmle3(or_spec, data=df, node_list, learner_list)

# parse tmle3_fit into AIPW_tmle class
AIPW_tmle$
  new(A=df$exposure,Y=df$outcome,tmle_fit = tmle3_fit,verbose = TRUE)$
  summary()
```





