---
title: "Getting Started with AIPW"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with AIPW}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  comment = "#>"#,
  # cache=TRUE
)
```


Contents:

  * [Installation](#Installation)
  * [One-line version](#one_line)
  * [Longer version](#details)
    + [Create an AIPW object](#constructor)
    + [Fit the object](#fit)
    + [Calculate average treatment effects](#ate)
  * [Parallelization](#par)
  * [Using tmle/tmle3 as input](#tmle_input)
    + [tmle](#tmle)
    + [tmle3](#tmle3)
    + [Compare AIPW estimates from different input methods](#compare)


## <a id="Installation"></a>Installation
1. Install AIPW from [GitHub](https://github.com/yqzhong7/AIPW)

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("yqzhong7/AIPW")
```

2. Install [SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/vignettes/Guide-to-SuperLearner.html) or [sl3](https://tlverse.org/sl3/articles/intro_sl3.html))

```{r, eval = FALSE}
#SuperLearner
install.packages("SuperLearner")
#sl3
remotes::install_github("tlverse/sl3")
```

## Input data for analyses 

```{r example data}
library(AIPW)
library(SuperLearner)
library(ggplot2)
set.seed(123)
data("eager_sim_obs")
cov = c("eligibility","loss_num","age", "time_try_pregnant","BMI","meanAP")
```

## Using AIPW to estimate the average treatment effect
### <a id="one_line"></a>One line version (Method chaining from R6class)
Using native AIPW class allows users to define different covariate sets for the exposure and the outcome models, respectively.

```{r one_line}
AIPW_SL <- AIPW$new(Y= eager_sim_obs$sim_Y,
                    A= eager_sim_obs$sim_A,
                    W= subset(eager_sim_obs,select=cov), 
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 10,
                    verbose=FALSE)$
  fit()$
  summary(g.bound = 0.25)$
  plot.p_score()
```

### <a id="details"></a>A more detailed tutorial

#### 1. <a id="constructor"></a>Create an AIPW object 

  * ##### Use [SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/index.html) libraries
```{r SuperLearner, message=FALSE,eval=F}
library(AIPW)
library(SuperLearner)

#SuperLearner libraries for outcome (Q) and exposure models (g)
sl.lib <- c("SL.mean","SL.glm")

#construct an aipw object for later estimations 
AIPW_SL <- AIPW$new(Y= eager_sim_obs$sim_Y,
                    A= eager_sim_obs$sim_A,
                    W= subset(eager_sim_obs,select=cov), 
                    Q.SL.library = sl.lib,
                    g.SL.library = sl.lib,
                    k_split = 10,
                    verbose=FALSE)
```

  * ##### Use [sl3](https://tlverse.org/sl3/index.html) libraries 
Metalearner is required to combine the estimates from stacklearner!
```{r sl3, eval=F}
library(AIPW)
library(sl3)

##construct sl3 learners for outcome (Q) and exposure models (g)
lrnr_glm <- Lrnr_glm$new()
lrnr_mean <- Lrnr_mean$new()
#stacking two learner (this will yield estimates for each learner)
stacklearner <- Stack$new(lrnr_glm, lrnr_mean) 
#metalearner is required to combine the estimates from stacklearner
metalearner <- Lrnr_nnls$new()
sl3.lib <- Lrnr_sl$new(learners = stacklearner,
                       metalearner = metalearner)

#construct an aipw object for later estimations 
AIPW_sl3 <- AIPW$new(Y= eager_sim_obs$sim_Y,
                     A= eager_sim_obs$sim_A,
                     W= subset(eager_sim_obs,select=cov), 
                     Q.SL.library = sl3.lib,
                     g.SL.library = sl3.lib,
                     k_split = 10,
                     verbose=FALSE)
```

#### 2. <a id="fit"></a>Fit the AIPW object

This step will fit the data stored in the AIPW object to obtain estimates for later average treatment effect calculations.

```{r}
#fit the AIPW_SL object
AIPW_SL$fit()
```

#### 3. <a id="ate"></a>Calculate average treatment effects 

  * ##### Estimate the ATE with propensity scores truncation
```{r}
#estimate the average causal effects from the fitted AIPW_SL object 
AIPW_SL$summary(g.bound = 0.25) #propensity score truncation 
```

  * ##### Check the balance of propensity scores by exposure status after truncation
```{r ps_trunc}
library(ggplot2)
AIPW_SL$plot.p_score()
```

## <a id="par"></a>Parallelization with future.apply

In default setting, the `AIPW$fit()` method will be run sequentially. The current version of AIPW package supports parallel processing implemented by [future.apply](https://github.com/HenrikBengtsson/future.apply) package under the [future](https://github.com/HenrikBengtsson/future) framework. Before creating a `AIPW` object, simply use `future::plan()` to enable parallelization and `set.seed()` to take care of the random number generation (RNG) problem:

```{r parallel, eval=FALSE}
# install.packages("future.apply")
library(future.apply)
plan(multiprocess, workers=2, gc=T)
set.seed(888)
AIPW_SL <- AIPW$new(Y= eager_sim_obs$sim_Y,
                    A= eager_sim_obs$sim_A,
                    W= subset(eager_sim_obs,select=cov), 
                    Q.SL.library = sl3.lib,
                    g.SL.library = sl3.lib,
                    k_split = 10,
                    verbose=FALSE)$fit()$summary()
```



## <a id="tmle_input"></a>Use `tmle`/`tmle3` fitted object as input

AIPW shares similar intermediate estimates with the Targeted Maximum Likelihood / Minimum Loss-Based Estimation (TMLE). Therefore, `AIPW_tmle` class is designed for using `tmle`/`tmle3` fitted object as input. Details about these two packages can be found [here](https://www.jstatsoft.org/article/view/v051i13) and [here](https://tlverse.org/tlverse-handbook/)

#### 1. <a id="tmle"></a>`tmle`

As shown in the message, [tmle](https://cran.r-project.org/web/packages/tmle/index.html) only support sample splitting in the outcome model. 

```{r tmle}
library(tmle)
library(SuperLearner)

tmle_fit <- tmle(Y=eager_sim_obs$sim_Y,
                 A=eager_sim_obs$sim_A,
                 W=subset(eager_sim_obs,select=cov), 
                 Q.SL.library=c("SL.mean","SL.glm"),
                 g.SL.library=c("SL.mean","SL.glm"),
                 family="binomial",
                 cvQinit = TRUE)

cat("\nEstimates from TMLE\n")
unlist(tmle_fit$estimates$ATE)
unlist(tmle_fit$estimates$RR)
unlist(tmle_fit$estimates$OR)

cat("\nEstimates from AIPW\n")
a_tmle <- AIPW_tmle$
  new(A=eager_sim_obs$sim_A,Y=eager_sim_obs$sim_Y,tmle_fit = tmle_fit,verbose = TRUE)$
  summary(g.bound=0.025)
```

#### 2. <a id="tmle3"></a> `tmle3`

Notably, [tmle3](https://github.com/tlverse/tmle3) conducts sample splitting and propensity truncation (0.025) by default.

```{r tmle3}
library(tmle3,quietly = TRUE)
library(sl3,quietly = TRUE)

node_list <- list(A = "sim_A",Y = "sim_Y",W = colnames(eager_sim_obs)[-1:-2])
or_spec <- tmle_OR(baseline_level = "0",contrast_level = "1")
tmle_task <- or_spec$make_tmle_task(eager_sim_obs,node_list)
lrnr_glm <- make_learner(Lrnr_glm)
lrnr_mean <- make_learner(Lrnr_mean)
sl <- Lrnr_sl$new(learners = list(lrnr_glm,lrnr_mean))
learner_list <- list(A = sl, Y = sl)
tmle3_fit <- tmle3(or_spec, data=eager_sim_obs, node_list, learner_list)

cat("\nEstimates from TMLE\n")
tmle3_fit$summary

# parse tmle3_fit into AIPW_tmle class
cat("\nEstimates from AIPW\n")
a_tmle3<- AIPW_tmle$
  new(A=eager_sim_obs$sim_A,Y=eager_sim_obs$sim_Y,tmle_fit = tmle3_fit,verbose = TRUE)$
  summary(g.bound=0)
```


#### 3. <a id="compare"></a> Compare the AIPW estimates from different input methods

```{r compare, echo=F, fig.width=12, collapse=FALSE}
options(width = 12)
library(knitr)
cat("True Estimates")
true_est <- rbind(attributes(eager_sim_obs)$true_rd, exp(attributes(eager_sim_obs)$true_logrr), exp(attributes(eager_sim_obs)$true_logor))
row.names(true_est) <- row.names(AIPW_SL$result)
colnames(true_est) <- "True Estimate"
kable(true_est,digits=3)
      
cat("Native AIPW function with sample splitting = 10")
kable(AIPW_SL$result,digits=3)

cat("From tmle fit object with sample splitting = 10 in the outcome model only")
kable(a_tmle$result,digits=3) 

cat("From tmle3 fit object with sample splitting = 10 (tmle3 default)")
kable(a_tmle3$result,digits=3)
```

