---
title: "Getting Started with AIPW"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with AIPW}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  comment = "#>"
)
```

-------

Contents:

  * [Installation](#Installation)
  * [One-line version](#one_line)
  * [Longer version](#details)
    + [Create an AIPW object](#constructor)
    + [Fit the object](#fit)
    + [Calculate average treatment effects](#ate)


## <a id="Installation"></a>Installation
1. Install AIPW from [GitHub](https://github.com/yqzhong7/AIPW)

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("yqzhong7/AIPW")
```

2. Install [SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/vignettes/Guide-to-SuperLearner.html) or [sl3](https://tlverse.org/sl3/articles/intro_sl3.html))

```{r, eval = FALSE}
#SuperLearner
install.packages("SuperLearner")
#sl3
remotes::install_github("tlverse/sl3")
```

## Input data for analyses

```{r example data}
set.seed(888)
N <- 200
outcome <- rbinom(N,1,0.3)
exposure <- rbinom(N,1,0.2)
#covaraites for outcome model (Q)
covariates.Q <- matrix(c(rbinom(N,1,0.4),
                                rnorm(N,mean = 0,sd=1),
                                rpois(N,lambda = 2)),
                              ncol=3)
#covariates for exposure model (g)
covariates.g <- matrix(c(rbinom(N,1,0.4),
                                rnorm(N,mean = 0,sd=1),
                                rpois(N,lambda = 2)),
                              ncol=3)

# covariates.g <- c(rbinom(N,1,0.4)) #a vector of a single covariate is also supported
```

## Using AIPW to estimate the average treatment effect
### <a id="one_line"></a>One line version (Method chaining from R6class)

```{r one_line}
library(AIPW)
library(SuperLearner)
library(ggplot2)
AIPW_SL <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = c("SL.mean","SL.glm"),
                    g.SL.library = c("SL.mean","SL.glm"),
                    k_split = 3,
                    verbose=FALSE)$
  fit()$
  calculate_result(g.bound = 0.25)$
  plot.p_score()
```

### <a id="details"></a>A more detailed tutorial

#### 1. <a id="constructor"></a>Create an AIPW object 

  * ##### Use [SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/index.html) libraries
```{r SuperLearner, message=FALSE}
library(AIPW)
library(SuperLearner)

#SuperLearner libraries for outcome (Q) and exposure models (g)
sl.lib <- c("SL.mean","SL.glm")

#construct an aipw object for later estimations 
AIPW_SL <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = sl.lib,
                    g.SL.library = sl.lib,
                    k_split = 3,
                    verbose=FALSE)
```

  * ##### Use [sl3](https://tlverse.org/sl3/index.html) libraries 
Metalearner is required to combine the estimates from stacklearner!
```{r sl3}
library(AIPW)
library(sl3)

##construct sl3 learners for outcome (Q) and exposure models (g)
lrnr_glm <- Lrnr_glm$new()
lrnr_mean <- Lrnr_mean$new()
#stacking two learner (this will yield estimates for each learner)
stacklearner <- Stack$new(lrnr_glm, lrnr_mean) 
#metalearner is required to combine the estimates from stacklearner
metalearner <- Lrnr_nnls$new()
sl3.lib <- Lrnr_sl$new(learners = stacklearner,
                      metalearner = metalearner)

#construct an aipw object for later estimations 
AIPW_sl3 <- AIPW$new(Y= outcome,
                    A= exposure,
                    W.Q=covariates.Q, 
                    W.g=covariates.g,
                    Q.SL.library = sl3.lib,
                    g.SL.library = sl3.lib,
                    k_split = 3,
                    verbose=FALSE)
```

#### 2. <a id="fit"></a>Fit the AIPW object

This step will fit the data stored in the AIPW object to obtain estimates for later average treatment effect calculations.

```{r}
#fit the AIPW_SL object
AIPW_SL$fit()
```

#### 3. <a id="ate"></a>Calculate average treatment effects 

  * ##### Check the balance of propensity scores by exposure status before ATE calculation
```{r ps_raw, fig.width=4}
library(ggplot2)
AIPW_SL$plot.p_score()
```

  * ##### Estimate the ATE with propensity scores truncation
```{r}
#estimate the average causal effects from the fitted AIPW_SL object 
AIPW_SL$calculate_result(g.bound = 0.25) #propensity score truncation 
```

  * ##### Check the balance of propensity scores by exposure status after truncation
```{r ps_trunc}
AIPW_SL$plot.p_score()
```
